Notes of Quantum Field Theory in a Nutshell

### 2016-08-13 QFT/Notes of QFT in a Nutshell/Notes of QFT in a Nutshell.tex
New complex geometry note.
Update QFT in nutshell's notes.

\chapter{Part I: Motivation and Foundations}
\label{sec:QFT_in_a_Nutshell.Part_I}
\section{I:2 Path Integral Formulation of Quantum Physics}
\label{sec:QFT_in_a_Nutshell_Part_I.2}
Here the path integral formulation of quantum mechanics is introduced.
The intuition comes from a limiting case of the traditional double 
slit electron interference experiment. (\textbf{pp.7 to 10}) Then 
it calculates the transition probability $\braket{q_F|e^{-iHT}|q_I}$
by divide it into a infinite of steps:
\begin{align}
    \braket{q_F|e^{-iHT}|q_I} = \lim_{N\to \infty}
    \braket{q_F|e^{iH\delta t}e^{iH\delta t}\cdots e^{iH\delta t}|q_I}
    (\text{with }N\delta t = T)
\end{align}

@@

\section{I:3 From Mattress to Field}
\label{sec:QFT_in_a_Nutshell_Part_I.3}
\paragraph{Note} After reading this section and several comments online,
I realize that this book put more emphasize on physical
intuition than on mathematical rigor. To illustrate this point, I
quote a sentence from A. Zee's homepage at UCSB, which comes from
a review of this book (QFT in a Nutshell):

@@

Next, we evaluate the disturbed case of free field theory:
\begin{align}
    Z = \int D\phi e^{i\int d^4x \left( \frac{1}{2} 
        [(\partial\phi)^2-m^2\phi^2]+J\phi \right)
        }
\end{align}
The integration is rather lengthy. It involves a trick called
"to imagine discretizing spacetime". In my opinion, such a trick
is to the best, a fancy intuition. There maybe mathematical proof,
but is neglected by the author. However, the result is surprisingly
rich in intuition and analogy. It start with the discretized
analogy with \eqref{eq:QFT_in_a_Nutshell_Part_I.2.appdx1.1}. One
is gradually lead to believe that the result of:
\begin{align}
    Z = \int D\phi e^{ i\int d^4x 
        \left(-\frac{1}{2}(\partial^2+m^2)\phi+J\phi \right)}
\end{align}
is:
\begin{align}
    Z(J) = \mathcal{C}e^{-(i/2) \int\int d^4xd^4y J(x)D(x-y)J(y)}
\end{align}
where $\mathcal{C}$ is just some factor which will proven to be
unimportant. $D(x-y)$ is the solution of:
\begin{align}
    \label{eq:QFT_in_a_Nutshell_Part_I.3.2}
    -(\partial^2+m^2)D(x-y)=\delta^{(4)}(x-y)
\end{align}

@@

One check by direct substitution that:
\begin{align}
    D(x-y) = \int \frac{d^4k}{(2\pi)^4} 
        \frac{e^{ik(x-y)}}{k^2-m^2+i\varepsilon}
\end{align}
is the solution to \ref{eq:QFT_in_a_Nutshell_Part_I.3.2}.

Then the author tries to explain how to evaluate $D(x)$, which
I fail to reproduce. The final result is equation (23) on page 
24. He notes that
\begin{quote}
    Physically, $D(x)$ describes the amplitude for a disturbance 
    in the field to propagate from the origin to $x$. 
\end{quote}
In different region of spacetime, the value of $D(x)$ varies greatly.
When $x$ is in timelink region, $D(x)$ is the superposition of
plane waves, while the phase is opposite between $x^0=t<0$ and
$x^0=t>0$. When $x$ is in the spacelike region, the $D(x)$ exhibits
a strange behavior that the particle can leak outside lightcone,
over a distance of order $m^{-1}$. This is consistent with the
Heisenberg's uncertainty principle.

@@

\textbf{Unclear points}: I cannot follow all the calculation for 
$D(x)$, because my result is different from the book's. Also, the
physics picture about $D(x)$ on page 24, is badly reasoned. The
argument is tenuous, in my opinion.
\chapter{Bibliography}
\begin{thebibliography}{1}
	\bibitem{zee} A. Zee. Quantum Field Theory in a Nutshell
		2ed. PUP.
\end{thebibliography}
\chapter{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
	Commons Attribution-NonCommercial-ShareAlike 4.0 International 
	License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.

>>>>

\chapter{Part I: Motivation and Foundations}
\label{sec:QFT_in_a_Nutshell.Part_I}
\section{I.2 Path Integral Formulation of Quantum Physics}
\label{sec:QFT_in_a_Nutshell_Part_I.2}
Here the path integral formulation of quantum mechanics is introduced.
The intuition comes from a limiting case of the traditional double 
slit electron interference experiment. (\textbf{pp.7 to 10}) Then 
it calculates the transition probability $\braket{q_F|e^{-iHT}|q_I}$
by divide it into a infinite of steps:
\begin{align}
    \braket{q_F|e^{-iHT}|q_I} = \lim_{N\to \infty}
    \braket{q_F|e^{iH\delta t}e^{iH\delta t}\cdots e^{iH\delta t}|q_I}
    (\text{with }N\delta t = T)
\end{align}

@@

\section{I.3 From Mattress to Field}
\label{sec:QFT_in_a_Nutshell_Part_I.3}
\paragraph{Note} After reading this section and several comments online,
I realize that this book put more emphasize on physical
intuition than on mathematical rigor. To illustrate this point, I
quote a sentence from A. Zee's homepage at UCSB, which comes from
a review of this book (QFT in a Nutshell):

@@

Next, we evaluate the disturbed case of free field theory:
\begin{align}
    Z = \int D\phi e^{i\int d^4x \left( \frac{1}{2} 
        [(\partial\phi)^2-m^2\phi^2]+J\phi \right)
        }
\end{align}
The integration is rather lengthy. It involves a trick called
"to imagine discretizing spacetime". In my opinion, such a trick
is to the best, a fancy intuition. There maybe mathematical proof,
but is neglected by the author. However, the result is surprisingly
rich in intuition and analogy. It start with the discretized
analogy with \eqref{eq:QFT_in_a_Nutshell_Part_I.2.appdx1.1}. One
is gradually lead to believe that the result of:
\begin{align}
    Z = \int D\phi e^{ i\int d^4x 
        \left(-\frac{1}{2}(\partial^2+m^2)\phi+J\phi \right)}
\end{align}
is:
\begin{align}
    \label{eq:QFT_in_a_Nutshell_Part_I.3.3}
    Z(J) = \mathcal{C}e^{-(i/2) \int\int d^4xd^4y J(x)D(x-y)J(y)}
\end{align}
where $\mathcal{C}$ is just some factor which will proven to be
unimportant. $D(x-y)$ is the solution of:
\begin{align}
    \label{eq:QFT_in_a_Nutshell_Part_I.3.2}
    -(\partial^2+m^2)D(x-y)=\delta^{(4)}(x-y)
\end{align}

@@

One checks by direct substitution that:
\begin{align}
    D(x-y) = \int \frac{d^4k}{(2\pi)^4} 
        \frac{e^{ik(x-y)}}{k^2-m^2+i\varepsilon}
\end{align}
is the solution to \ref{eq:QFT_in_a_Nutshell_Part_I.3.2}.

We also denote the exponent in \ref{eq:QFT_in_a_Nutshell_Part_I.3.3}
as $W(J)$. So:
\begin{align}
    \label{eq:}
    W(J) = -\frac{1}{2} \int\int d^4x d^4y J(x)D(x-y)J(y) \\
    Z= \mathcal{C}e^{iW(J)}
\end{align}

\paragraph{Free propagator}

The function $D(x)$ is also called a propagator. It "plays an 
essential role in quantum field theory. As the inverse of a
differential operator it is clearly closely related to the Green’s
function you encountered in a course on electromagnetism".

Then the author tries to explain how to evaluate $D(x)$, which
I fail to reproduce. The final result is equation (23) on page 
24:
\begin{align}
    D(x)= -i\int \frac{d^3k}{(2\pi)^3 2\omega_k}
    \left( e^{-i(w_k t-\vec{k}\cdot\vec{x})} \theta(x^0)
        + e^{i(w_k t-\vec{k}\cdot\vec{x})} \theta(-x^0)
    \right)
\end{align}

He notes that
\begin{quote}
    Physically, $D(x)$ describes the amplitude for a disturbance 
    in the field to propagate from the origin to $x$. 
\end{quote}
In different region of spacetime, the value of $D(x)$ varies greatly.
When $x$ is in timelink region, $D(x)$ is the superposition of
plane waves, while the phase is opposite between $x^0=t<0$ and
$x^0=t>0$. When $x$ is in the spacelike region, the $D(x)$ exhibits
a strange behavior that the particle can leak outside lightcone,
over a distance of order $m^{-1}$. This is consistent with the
Heisenberg's uncertainty principle.

@@

\textbf{Unclear points}: I cannot follow all the calculation for 
$D(x)$, because my result is different from the book's. Also, the
physics picture about $D(x)$ on page 24, is badly reasoned. The
argument is tenuous, in my opinion.

\section{I.4 From Field to Particle to Force}
\label{sec:I.4 From Field to Particle to Force}

\textbf{Note}: discussion in this section is based on the \textit{free
field theory}.

\paragraph{From field to particle}

With the Fourier transformation, several physical properties
could be extrated from $W(J)$. The Fourier transformed $W(J)$ is:
\begin{align}
    W(J) = -\frac{1}{2}\int \frac{d^4}{(2\pi)^4}J(k)^*
        \frac{1}{k^2-m^2+i\epsilon}J(k)
\end{align}
If separating $J$ into $J(x)=J_1(x)+J_2(x)$, then naturally, the
Fourier transformed calculation will include terms of the form:
$J^*_1J_1,J^*_1J_2,J^*_2J_1,J^*_2J_2$. We see that the denominator
in $W(J)$ contains $k^2-m^2$. Somehow the author wants to argue that
the term $J^*_2J_1$ describes the influence of a particle of mass $m$
propagating from the source $1$ to the sink $2$. But his language is loose.

This part also introduces the jargon \textit{on mass shell} and the
language of "virtual particles" moving in $k$-space.

\paragraph{From particle to force}
With two delta source $J_a(x)=\delta^(3)(\vec{x}-\vec{x}_a)$, one can
describe two massive lumps setting on $\vec{x}_1$ and $\vec{x}_2$ and
not moving at all. These two lumps, coupled with the free field $\phi$,
will attract each other. To show the attraction, the author calculates
explicitly the $W(J)$, and identifies $iW(J) \equiv -iET$ (the
identification is substantiate by $e^{-iET} = Z = e^{iW(J)}$).
Finally, it shows:
\begin{align}
    E= -\frac{1}{4\pi r}e^{-mr}
\end{align}
This is actually the Yukawa model for the attraction between nucleons
in the atomic nucleus. A plot is in due:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/Yukawa-potential.pdf}
    \caption{Plot of Yukawa potential $-\frac{e^{-r}}{4 \pi  r}$}
\end{figure}

It shows us several information:
\begin{itemize}
    \item In the long distance, the potential drops off over the
        distance scale of $1/m$.
    \item In the short range, the force is $1/r^2$.
    \item The energy gets smaller when the two sources come closer.
        It tends to $-\infty$ as $r\to 0$.
\end{itemize}

\paragraph{Origin of force}
% TODO stops here.
\chapter{Bibliography}
\begin{thebibliography}{1}
	\bibitem{zee} A. Zee. Quantum Field Theory in a Nutshell
		2ed. PUP.
\end{thebibliography}
\chapter{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
	Commons Attribution-NonCommercial-ShareAlike 4.0 International 
	License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
